### 2017-05-04 gpu-05-03-accelerator.tex
Start working on integration: block out double algorithm, unify symbol usage

\subsection{Analysis of CPU Parallelism}

@@

\begin{table}[b]	% [htb]
% \center\begin{tabular}{l c c c c c c c c d{3.1}}
% Device & Architecture & cores & Clock & \multicolumn{2}{c}{GFlops} & L2 Cache & Memory & Bandwidth \\ 
%        &              &       & MHz   & single & double            & KB       & GB     & (GB/s) \\ 
% \hline
% K20m        & Tesla   &  2496 &  706        & 3520 & 1170 &      &  5 & 208 \\
% K40c        & Tesla   &  2880 &  745        & 4290 & 1430 &      & 12 & 208 \\
% GTX 980     & Maxwell &  2816 & 1126 (1216) & 4612 &  144 &      &  4 & 336.5 \\
% GTX Titan-X & Maxwell &  3072 & 1000 (1075) & 6144 &  192 & 2048 & 12 & 336.5 \\
%
%
%###########################################################################################
% SPECS FROM:
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-titan-x/specifications
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-980/specifications
% http://www.nvidia.com/content/PDF/kepler/Tesla-K40-Active-Board-Spec-BD-06949-001_v03.pdf
% http://www.nvidia.com/content/PDF/kepler/Tesla-K20-Active-BD-06499-001-v04.pdf
%
% FLOPS DATA:
% K40: http://international.download.nvidia.com/pdf/kepler/TeslaK80-datasheet.pdf
% K20 & K40: http://www.nvidia.com/content/tesla/pdf/nvidia-tesla-kepler-family-datasheet.pdf
% GTX980 & TITANX: https://en.wikipedia.org/wiki/List_of_Nvidia_graphics_processing_units
\center\begin{tabular}{lc c c c c}
                         && \multicolumn{4} {c}{Device} \\
\cline{3-6}
Specification            && K20m  & K40c  & GTX 980 & GTX Titan-X \\
\cline{1-1}\cline{3-6}
Number of Cores          && 2496  & 2880  & 2048    & 3072  \\
Clock (MHz)              && 706   & 745   & 1126    & 1000  \\
GFlops (single)          && 3520  & 4290  & 4612    & 6144  \\
GFlops (double)          && 1170  & 1430  & 144     & 192   \\
Memory (GB)              && 5     & 12    & 4       & 12    \\
Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5 \\
\cline{1-1}\cline{3-6}
\end{tabular}
\caption{Properties of the GPUs used in the evaluation}
\label{table-gpus}
\end{table}

\subsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on Nvidia's GTX TitanX,
GTX980, K40c and K20m.

@@

\subsection{Comparison of Compute Devices}
Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
GTX TitanX, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
sequential C++ version. These results are consistent with the relative
capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
the TitanX achieves the highest speedup of 86 relative to the baseline. At the
other end, the parallel CPU version attains a speedup factor of~20.9.

>>>>

\subsubsection{Analysis of CPU Parallelism}

@@

\begin{table}[b]	% [htb]
% \center\begin{tabular}{l c c c c c c c c d{3.1}}
% Device & Architecture & cores & Clock & \multicolumn{2}{c}{GFlops} & L2 Cache & Memory & Bandwidth \\ 
%        &              &       & MHz   & single & double            & KB       & GB     & (GB/s) \\ 
% \hline
% K20m        & Tesla   &  2496 &  706        & 3520 & 1170 &      &  5 & 208 \\
% K40c        & Tesla   &  2880 &  745        & 4290 & 1430 &      & 12 & 208 \\
% GTX 980     & Maxwell &  2816 & 1126 (1216) & 4612 &  144 &      &  4 & 336.5 \\
% GTX Titan-X & Maxwell &  3072 & 1000 (1075) & 6144 &  192 & 2048 & 12 & 336.5 \\
%
%
%###########################################################################################
% SPECS FROM:
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-titan-x/specifications
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-980/specifications
% http://www.nvidia.com/content/PDF/kepler/Tesla-K40-Active-Board-Spec-BD-06949-001_v03.pdf
% http://www.nvidia.com/content/PDF/kepler/Tesla-K20-Active-BD-06499-001-v04.pdf
%
% FLOPS DATA:
% K40: http://international.download.nvidia.com/pdf/kepler/TeslaK80-datasheet.pdf
% K20 & K40: http://www.nvidia.com/content/tesla/pdf/nvidia-tesla-kepler-family-datasheet.pdf
% GTX980 & TITANX: https://en.wikipedia.org/wiki/List_of_Nvidia_graphics_processing_units
\center\begin{tabular}{lc c c c c}
                         && \multicolumn{4} {c}{Device} \\
\cline{3-6}
Specification            && K20m  & K40c  & GTX 980 & GTX Titan-X \\
\cline{1-1}\cline{3-6}
Number of Cores          && 2496  & 2880  & 2048    & 3072  \\
Clock (MHz)              && 706   & 745   & 1126    & 1000  \\
GFlops (single)          && 3520  & 4290  & 4612    & 6144  \\
GFlops (double)          && 1170  & 1430  & 144     & 192   \\
Memory (GB)              && 5     & 12    & 4       & 12    \\
Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5 \\
\cline{1-1}\cline{3-6}
\\[-1.5ex]
\end{tabular}
\caption{Properties of the GPUs used in the evaluation}
\label{table-gpus}
\end{table}

\subsubsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on Nvidia's GTX TitanX,
GTX980, K40c and K20m.

@@

\subsubsection{Comparison of Compute Devices}
Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
GTX TitanX, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
sequential C++ version. These results are consistent with the relative
capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
the TitanX achieves the highest speedup of 86 relative to the baseline. At the
other end, the parallel CPU version attains a speedup factor of~20.9.

### 2017-05-04 gpu-05-05-perplexity.tex
Start working on integration: block out double algorithm, unify symbol usage

\subsection{Algorithm Convergence}

This section presents the evolution of the perplexity, i.e.\ the metric for
the solution quality, as a function of time for
two datasets, whose properties are given in Table~\ref{table-snap}. This
analysis shows that our implementations of the algorithm indeed achieve
convergence, and it gives an impression of the time required to achieve
convergence for the largest datasets and community sizes that fit into the memory
of a GPU and a CPU respectively. The dominant customer of memory is the
$\pi$ matrix, sized $N{\times}K$ for $K$ communities and $N$ vertices in the
graph. The runtime per iteration is independent of~$N$, but it depends on the
product of~$K$, the mini-batch size~$m$, and the neighbor sample size~$n$. On
the other hand, the number of iterations required to achieve sufficient
convergence does grow with the size of the graph.

@@

Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity development convergence for
com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
$m$=4096, $n$=32. Sufficient convergence is reached after 18~hours. Note
that this network graph is so large that the GTX Titan-X would be unable to
process more than $K$=256 communities.

>>>>

\subsubsection{Algorithm Convergence}

This section presents the evolution of the perplexity, i.e.\ the metric for
the solution quality, as a function of time for
two datasets, whose properties are given in Table~\ref{table-snap}. This
analysis shows that our implementations of the algorithm indeed achieve
convergence, and it gives an impression of the time required to achieve
convergence for the largest datasets and community sizes that fit into the memory
of a GPU and a CPU respectively. The dominant customer of memory is the
$\pi$ matrix, sized $N{\times}K$ for $K$ communities and $N$ vertices in the
graph. The runtime per iteration is independent of~$N$, but it depends on the
product of~$K$, the mini-batch size~$M$, and the neighbor sample size~|\Neighbors|. On
the other hand, the number of iterations required to achieve sufficient
convergence does grow with the size of the graph.

@@

Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity development convergence for
com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
$M$=4096, |\Neighbors|=32. Sufficient convergence is reached after 18~hours. Note
that this network graph is so large that the GTX Titan-X would be unable to
process more than $K$=256 communities.

### 2017-05-04 gpu-05-evaluation.tex
Place the chunks of text unmodified into one document
Start working on integration: block out double algorithm, unify symbol usage

\section{Evaluation}
\label{sec-evaluation}

@@

\begin{table}[tb]
\center\begin{tabular}{l r r c}
Network & \#Vertices & \#Edges \\
\hline
CA-HepPh        &    12,008 &    118,521 \\
com-DBLP        &   317,080 &  1,049,866 \\
% com-Youtube     & 1,134,890 &  2,987,624 \\
com-LiveJournal & 3,997,962 & 34,681,189 \\
\hline
\end{tabular}
\caption{Network graphs from the Stanford Snap collection}
\label{table-snap}
\end{table}

\input{05-02-sequential}
\input{05-03-accelerator}
\input{05-05-perplexity}

>>>>

\subsection{Evaluation}
\label{sec-evaluation}

@@

\begin{comment}
\begin{table}[tb]
\center\begin{tabular}{l r r c}
Network & \#Vertices & \#Edges \\
\hline
CA-HepPh        &    12,008 &    118,521 \\
com-DBLP        &   317,080 &  1,049,866 \\
% com-Youtube     & 1,134,890 &  2,987,624 \\
com-LiveJournal & 3,997,962 & 34,681,189 \\
\hline
\\[-1ex]
\end{tabular}
\caption{Network graphs from the Stanford Snap collection}
\label{table-snap}
\end{table}
\end{comment}

\begin{table*}
  \centering
  \begin{tabular}{l r r r l}
    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
                    &            &               & communities    &             \\
    \hline
    CA-HepPh        &    12,008  &    118,521    & n/a            & Paper citation network \\
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
    \hline
    \\[-1ex]
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{table-snap}
\end{table*}

\input{gpu-05-02-sequential}
\input{gpu-05-03-accelerator}
\input{gpu-05-05-perplexity}

### 2017-05-04 gpu-06-conclusion.tex
Start working on integration: block out double algorithm, unify symbol usage

\section{Conclusion and Future Work}
Modern machine learning algorithms are empowering their
users to solve or approximate solutions to previously intractable problems.
However, these advancements come at the cost of significant computational
complexity and long running learning processes which hinder our ability to reap
such algorithms' benefits when applied to large problems. There is a clear need
to assess such algorithms and identify optimization opportunities in order to
scale their performance.

@@

This paper focused on the acceleration of the SG-MCMC algorithm on a
single node. For future work, we are designing and implementing a distributed
version of the algorithm that uses the collective memory of a cluster, which will enable
processing of much larger datasets. These changes in context fundamentally
change the optimization landscape's boundaries and priorities.

>>>>

\section{Conclusion}
Modern machine learning algorithms are empowering their
users to solve or approximate solutions to previously intractable problems.
However, these advancements come at the cost of significant computational
complexity and long running learning processes which hinder our ability to reap
such algorithms' benefits when applied to large problems. There is a clear need
to assess such algorithms and identify optimization opportunities in order to
scale their performance.

@@
